---
title: "STAT 311: Problem Set 5 (R)"
author: "Your name here"
date: "03/16/2026"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message = F}
# Note: if you have never installed tidyverse, you will have to do so now. In the CONSOLE at the bottom, paste in and run the the following code (without the #):
# install.packages("tidyverse")

# You will see some text running for a bit, but eventually everything should be successful! Then run this chunk again.
library(tidyverse)
```

We will now simulate some German Tank data, obtain estimates of the number of tanks, and compare the estimators.

a)  Create three variables:
    -   `n_sim` for the number of simulations (set this to be at least 1000)

    -   `N_true` for the true number of tanks (make this 100); and

    -   `n` for the number of tanks captured (set this to 5)

```{r}
# code for part (a)

```

b)  Write a function called `sim_tanks` that takes in two arguments: the true number of tanks and the number of tanks captured. The function should simulate the German tank scenario by randomly sampling the tanks without replacement from the discrete uniform (use the `sample()`. function). 

The function should return a **vector** of length three of the estimates  under $\hat{N}_{0}$, $\hat{N}_{RB}$, and $\hat{N}_{URB}$, as defined on the homework.

```{r}
# code for part (b)

```

c)  Using the `replicate()` function, simulate this problem for `n_sim` iterations. Then use the `t()` function to take the *transpose* of the output such that you have 3 columns and `n_sim` rows. Be sure to store your output into a variable!

*In the Console, take a look at the output. Make sure you understand what each row/column represents!*

```{r}
set.seed(311)
# code for part (c)

```

d)  Using your simulations from (c), obtain the empirical MSE of $\hat{N}_{0}$, $\hat{N}_{RB}$, and $\hat{N}_{URB}$. You can access a specific column $j$ of matrix $M$ using the following code: `M[,j]`. Similarly, to access a specific row: `M[j,]`. Does what you find make sense? Briefly explain.

**Note**: do not use the empirical mean and variances to obtain the empirical MSE. Approximate the MSE of an estimator $\delta(X)$ from its definition of $E[(\delta(X) - \theta)^2]$. You should use the `mean()` function once per estimator.

```{r}
# three lines of code to approximate MSEs

```

**Answer: Replace with your answer.**

e)  Lastly, we will visualize the simulations of the estimates. Run the first line of code below and then take a look at what `df` looks like by typing `View(df)` in the Console.

Then fill in the remaining code and then set `eval = TRUE` before knitting. Briefly comment on what you notice.

```{r eval = FALSE}
df <- data.frame(___) |> # variable for your matrix holding estimates
  rename("X1" = 1, "RB" = 2, "RB_unbiased" = 3) |>
  pivot_longer(cols = all_of(1:3), names_to = "estimator", values_to = "estimate") 

df |>
  ggplot(aes(x = ______, # name of variable in df representing estimates
                fill = _____ ))+ # name of variable in df representing estimator
  geom_histogram(alpha = 0.3, binwidth = 2) +
  geom_vline(xintercept = ____ ) # variable representing true total number of tanks
```

**Answer: Replace with your answer.**
