---
title: "(Asymptotic) efficiency"
editor_options: 
  chunk_output_type: console
---

Let $X_{1},\ldots, X_{n} | \theta \overset{iid}{\sim} N(\theta, \sigma^2)$ where $\sigma^2$ known. We want to estimate $g(\theta) = \theta^2 + \sigma^2$. We have two candidate estimators:

$$\delta_{u}(\vec{X}) = \bar{X}^2 + \sigma^2(1 - 1/n) \qquad \text{ and }\quad \delta_{MLE}(\vec{X}) = \bar{X}^2 + \sigma^2$$
where $\delta_{u}$ is unbiased for $g(\theta)$ but the MLE is not so CRLB won't apply. We found that at small $n$, the unbiased  $\delta_{u}$ does not attain the CRLB. What happens as $n\rightarrow \infty$? 


```{r echo = F}
theta <- 1
s2 <- 25
```

Here, $\theta = `r theta`$ and $\sigma^2 = `r s2`$. We will visualize the MSE for both estimators at various values of $n$, and provide the CRLB for an unbiased estimator of $g(\theta)$.

```{r message = F, echo = F}
library(tidyverse)
n_vec <- 2:200
mse_mat <- matrix(NA, nrow = length(n_vec), ncol = 4)
for(i in 1:length(n_vec)){
  n <- n_vec[i]

  g_theta <- theta^2 + s2
  
  delta_u_var <- 4 * theta^2 * s2 / n + 2 * s2^2 / n^2
  delta_u_mse <- delta_u_var + 0
  
  mle_bias <- theta^2  + s2 * (1 + 1/n) - g_theta
  mle_var <- 4 * theta^2 * s2 / n + 2 * s2^2 / n^2
  mle_mse <- mle_var + mle_bias^2
  
  crlb <- 4 * theta^2 * s2 / n
  mse_mat[i,] <- c(delta_u_mse, mle_mse, crlb, n)
}
colnames(mse_mat) <- c("delta_u", "mle", "crlb", "n")
data.frame(mse_mat) |>
  pivot_longer(cols = c(1:2), names_to = "estimator", values_to = "mse") |>
  ggplot(aes(x = n, y = mse, col = estimator)) +
  geom_line(linewidth = 1) +
  geom_point(aes(x = n, y = crlb), size = 0.25, col = "#C77CFF") +
  scale_y_log10(n.breaks = 6) +
  labs(y = "MSE (log scale)",
       caption = "Dotted purple line represent CRLB at each n",
       subtitle = bquote(sigma^2 == .(s2))) +
  theme(text =element_text(size = 16))
```


What happens when we lower $\sigma^2$? Why?

```{r echo = F}
s2 <- 1
mse_mat <- matrix(NA, nrow = length(n_vec), ncol = 4)
for(i in 1:length(n_vec)){
  n <- n_vec[i]

  g_theta <- theta^2 + s2
  
  delta_u_var <- 4 * theta^2 * s2 / n + 2 * s2^2 / n^2
  delta_u_mse <- delta_u_var + 0
  
  mle_bias <- theta^2  + s2 * (1 + 1/n) - g_theta
  mle_var <- 4 * theta^2 * s2 / n + 2 * s2^2 / n^2
  mle_mse <- mle_var + mle_bias^2
  
  crlb <- 4 * theta^2 * s2 / n
  mse_mat[i,] <- c(delta_u_mse, mle_mse, crlb, n)
}
colnames(mse_mat) <- c("delta_u", "mle", "crlb", "n")
data.frame(mse_mat) |>
  pivot_longer(cols = c(1:2), names_to = "estimator", values_to = "mse") |>
  ggplot(aes(x = n, y = mse, col = estimator)) +
  geom_line(linewidth = 1) +
  geom_point(aes(x = n, y = crlb), size = 0.25, col = "#C77CFF") +
  scale_y_log10(n.breaks = 6) +
  labs(y = "MSE (log scale)",
       caption = "Dotted purple line represent CRLB at each n",
       subtitle = bquote(sigma^2 == .(s2)))+
  theme(text =element_text(size = 16))
```