---
title: "Size and level of a test example"
format: pdf
editor_options: 
  chunk_output_type: console
---
```{r message = F, echo = F}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F,
                      fig.width = 5, fig.height = 2.75, fig.align = "center")
```

\vspace{-3cm}
## Calculating size of a test 

Let $X_{1},\ldots, X_{100} | \theta \overset{iid}{\sim} \text{Bern}(\theta)$ where $\theta \in [0,1]$.
$$H_{0}: \theta \leq 0.5 \qquad \qquad H_{1}: \theta > 0.5$$
For our testing procedure $\delta$, define $r(\mathbf{X}) = \sum_{i=1}^{100} X_{i}$ and $R = \{r(\mathbf{X}) \geq 54 \}$.

### Visualize power function

We can see that the power function is increasing in $\theta$: $$\pi(\theta | \delta) = \text{Pr}\left(\sum_{i=1}^{100} X_{i} \geq 54 | \theta\right) = \sum_{k=54}^{100} \binom{100}{54}\theta^{54}(1-\theta)^{100-54}$$ 

```{r echo = F}
n <- 100
theta_seq <- seq(0,1,0.0001)
power <- 1 - pbinom(53, n, theta_seq)
data.frame(theta = theta_seq, power = power) |>
  ggplot(aes(x = theta, y = power)) +
  geom_line(linewidth = 1) +
  labs(x = expression(theta),
       y = expression(pi (theta ~ "|" ~ delta))) +
  theme(text =element_text(size = 15))
```

### Find size of test

How to find its size? Recall:
$$\alpha(\delta) = \sup_{\theta \in \Omega_{0}} \pi(\theta | \delta)$$

```{r echo = F}
size <- 1 - pbinom(53, n, 0.5)
size_df <- data.frame(x = c(0, 0.5),
                      xend = c(0.5, 0.5),
                      y = c(size, 0),
                      yend = c(size, size))
df <- data.frame(theta = theta_seq, power = power) |>
  mutate(space = ifelse(theta <= 0.5, "Omega_0", "Omega_1"))

ggplot() +
  geom_line(data = df, aes(x = theta, y = power, col = space),
            linewidth = 1) +
  #geom_segment(data = size_df, mapping = aes(x = x, xend = xend, y = y, yend = yend), linetype = "dashed") +
  # geom_hline(yintercept = size, linetype = "dashed") +
  labs(x = expression(theta),
       y = expression(pi (theta ~ "|" ~ delta))) +
  theme(text =element_text(size = 15)) +
  guides(col = "none")
```

We see that the sup occurs at $\theta = 0.5$! So:

$$\alpha(\delta) = \sup_{\theta \in \Omega_{0}} \text{Pr}\left(\sum_{i=1}^{100} X_{i} \geq 54 | \theta\right) =  \text{Pr}\left(\sum_{i=1}^{100} X_{i} \geq 54 | \theta = 0.5\right)$$
Using R, the size of our test is:

```{r}
# pbinom(x, n, p) = P(X <= x)
1 - pbinom(53, n, 0.5)
```

\clearpage
## Finding a level-$\alpha_{0}$ test

### Brute-force approach

Now suppose I consider tests $\delta_{c}$ that have rejection regions $R = \{r(\mathbf{X}) \geq c \}$ for some $c$, where $r(\mathbf{X}) = \sum_{i=1}^{100} X_{i}$ as before. I want to find a $c$ that yields a level-10 test.

The following plots $\text{Pr}(\sum X_{i} \geq c | \theta = 0.5)$ for $c = 0,1,\ldots, 100$. How can we use this to determine which $c$ gives us a level-0.10 test?

```{r message=F, echo = F, fig.height=3.25}
# option 1
theta <- 0.5

c_seq <- 0:100
alpha0 <- 0.10
power_c <- 1 - pbinom(c_seq - 1, n, theta)
data.frame(c = c_seq, power = power_c) |>
  mutate(below = power <= alpha0) |>
  ggplot(aes(x = c, y = power)) +
  geom_point() +
    theme(text =element_text(size = 15)) +
  geom_hline(yintercept = alpha0, linetype = "dashed",
             linewidth = 0.35) +
  scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(n.breaks = 10) +
   guides(col = "none") +
  labs(y = expression(pi(theta ~ "=" ~  0.5 ~ "|" ~ delta[c])))
c_star <- c_seq[min(which(power_c < alpha0))]
```

```{r echo = F, eval = F}
data.frame(c = c_seq, power = power_c) |>
  mutate(below = c == c_star) |>
  ggplot(aes(x = c, y = power, col = below)) +
  geom_point() +
    theme(text =element_text(size = 15)) +
  # geom_hline(yintercept = alpha0, linetype = "dashed") +
  scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(n.breaks = 10) +
   guides(col = "none") +
  scale_color_manual(values = c("black", "magenta")) +
  labs(y = expression(pi(theta ~ "=" ~  0.5 ~ "|" ~ delta[c])))
```

We can see that these probabilities drop below 0.10 when $c = `r c_star`$! So the testing procedure that has rejection region $R = \{r(\mathbf{X}) \geq `r c_star` \}$ is a level-0.10 test of these hypotheses!

```{r}
# Confirm: P(sum Xi >= 57 | theta = 0.5) <= 0.10
1 - pbinom(57 - 1, n, 0.5)
```

\textbf{Question:} does this choice of $c = `r c_star`$ yield the only level-0.10 test?

### CLT approximation

Using the CLT approach, we found 
$$\text{Pr}\left(\sum_{i=1}^{100} X_{i} \geq c | \theta = 0.5\right) \approx 1- \Phi\left(\frac{c - 50}{5} \right)$$
We will also plot this as a function of $c$ to find a level-0.10 test, next to the plot from before with the exact probabilites shown on the previous plot. Note that they are quite similar! 

```{r echo = F, fig.height=3.25}
data.frame(c = c_seq, Exact = power_c, CLT = 1 - pnorm( (c_seq - 50)/5)) |>
  pivot_longer(cols = 2:3, names_to = "method", values_to = "prob") |>
  ggplot(aes(x = c, y = prob)) +
  geom_point(size = 0.65) +
  geom_hline(yintercept = alpha0, linetype = "dashed",
             linewidth = 0.35)+
  theme(text =element_text(size = 13)) +
    scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(n.breaks = 10) +
  labs(y = "Probability") +
  # labs(y = expression("CLT approx. of " ~ pi(theta ~ "=" ~  0.5 ~ "|" ~ delta[c]))) +
  facet_wrap(~method)
```

We can see that the CLT yields the same choice of $c = `r c_star`$. 

Let's obtain the CLT approximated probability:

```{r}
# verify CLT approximation
1 - pnorm( (57 - 50) / 5)
```
