---
title: "LRT asymptotics example"
---

## Set-up

Recall the coin-flipping scenario again! We have $X_{1},\ldots, X_{n}$ are a random sample from a $\text{Bern}(\theta)$ distribution, where $\theta$ is the probability of Heads. We had the hypotheses

$$H_{0}: \theta = \theta_{0} \qquad \text{vs.} \qquad H_{1}: \theta \neq \theta_{0}$$ We can conduct a LRT of these hypotheses. We found the likelihood ratio statistic to be:

$$\Lambda(\mathbf{X}) = \left(\frac{n\theta_{0}}{\sum X_i}\right)^{\sum X_{i}} \left( \frac{n(1-\theta_{0})}{n - \sum X_i}\right)^{n - \sum X_i}$$The LRT says reject when $\Lambda(\mathbf{x}) \leq k$ for some $k \in [0,1]$.

Consider $\theta_{0} = 0.6$ and $n = 60$ coin flips. Suppose we observed $\sum X_i = 20$ heads. Should we reject $H_{0}$? Let's obtain a p-value!

### Approximate $p$-value

We know that under $H_{0}$, in this problem the random variable $-2\log \Lambda(\mathbf{X}) \overset{\cdot}{\sim} \chi^2_{1}$ as $n \rightarrow \infty$. Let's obtain the approximate p-value.

```{r}
n <- 60
theta0 <- 0.6
y_obs <- 20
Lambda_obs <- ((n*theta0/y_obs)^y_obs) * (n*(1-theta0)/(n - y_obs))^(n - y_obs)
Lambda_obs
pval_approx <- 1 - pchisq(-2*log(Lambda_obs), df = 1)
pval_approx
```

### Exact $p$-value

Let's compare the approximation to the exact p-value under this LRT.

```{r echo = F}
y <- 0:n
prob_y <- dbinom(0:n, n, theta0)
Lambda <- ((n*theta0/y)^y) * (n*(1-theta0)/(n - y))^(n - y)
Lambda_sort <- sort(Lambda)
prob_y_sort <- prob_y[sort(Lambda, index.return = T)$ix]
y_sort <- (0:n)[sort(Lambda, index.return = T)$ix]
y <- 20
Lambda_obs <- ((n*theta0/y)^y) * (n*(1-theta0)/(n - y))^(n - y)
pval_exact <- sum(prob_y_sort[which(Lambda_sort <= Lambda_obs)])
# y_sort[which(Lambda_sort <= Lambda_obs)]
```

The exact p-value for this data is (I challenge you to try obtaining this on your own) is `r pval_exact`.

Notice how similar these values are! Relies on large $n$.
