---
title: "Method of Moments: Mark-Recapture Simulation"
date: "2024-03-05"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

### Set-up

Recall the following scenario: we would like estimate the number of individuals $N$ of a species in a particular location. The mark-recapture sampling scheme proceeds as follows:

1.  Sample $r$ individuals from the population and mark/tag them all. Release the individuals and wait some time.
2.  After waiting, take a second sample of $m$ individuals from the population. Count how many of the $m$ are marked.

Defining $X$ as the number of individuals in the second sample who were tagged, we obtained the following method of moments estimator: $\hat{N}_{MM} = \frac{rm}{X}$.

### Simulation

Let's simulate data to see how well this estimator performs. Note that in simulating data, I get to choose/know the *true* value of $N$.

```{r}
# set true value
N_true <- 2000
```

Now, I will determine how many to take in the first sample. Then I create a vector of 1's and 0's called `marked`, where I have $r$ 1's representing all the tagged individuals, and the remaining $N-r$ 0's representing all the untagged.

```{r}
# number in first sample
r <- 200

# make vector of "marked" and "unmarked" individuals after initial sample
marked <- c(rep(1, r), rep(0, N_true -r))
```

We write a function that performs one iteration of the second sample: sampling $m$ individuals without replacement from the population again, and counting how many were `marked`.

```{r}
# write function to simulate second sample
mark_recapture <- function(marked, m){
  samp2 <- sample(marked, size = m, replace = F)
  x <- sum(samp2)
  return(x)
}
```

Now we actually simulate! I can choose how many simulations to run. Then I use the `replicate()` function to repeatedly perform the simulations. The outcome all the replicates gets stored in a vector of length `n_sims` called `R`, which I use to obtain the method of moments estimates over repeated samples.

```{r}
# choose number of simulations
n_sims <- 10000
# number in second sample
m <- 300

# simulate!
R <- replicate(n_sims, mark_recapture(marked, m))

# calculate MoM estimate
N_hat_vec <- r * m / R

# plot
library(tidyverse)
data.frame(N_hat = N_hat_vec) %>%
  ggplot(., aes(x = N_hat)) +
  geom_histogram() +
  geom_vline(xintercept = N_true, col = "red", linetype = "dashed")
```
