---
title: "Power functions"
author: "Becky Tang"
date: "2024-04-16"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F,
                      fig.width = 6, fig.height = 4)
```

## Plotting power function

Recall the following statistical model and hypotheses:

$$X_{1},\ldots, X_{10} |\theta \sim \text{Bern}(\theta), \qquad \theta \in [0,1]$$ $$H_{0}: \theta = \frac{1}{2} \qquad \text{vs.} \qquad H_{1}: \theta \neq \frac{1}{2}$$ We have the following testing procedure $\delta_{1}$: $r(\mathbf{X}) = |\sum_{i=1}^{10} X_{i} - 5|, \text{ with rejection region } R = \{r(\mathbf{X}) = 5 \}$

So the power function is $\pi(\theta | \delta_{1}) = P( \sum_{i=1}^{10} X_{i} \in \{0, 10\}|\theta)$.

```{r}
power_fn <- function(theta, n, r_vals){
  # theta is possible vector of theta_vals
  # n = number of trials
  # r_vals = values in rejection region
  n_theta <- length(theta)
  ret <- rep(NA, n_theta)
  for(i in 1:n_theta){
    ret[i] <- sum(dbinom(r_vals, n, theta[i]))
  }
  return(ret)
}


# create sequence of theta values for plotting purposes
theta_vals <- seq(0, 1, 0.01)
n <- 10
delta1 <- power_fn(theta_vals, n, c(0, 10))
```

```{r}
library(tidyverse)
data.frame(theta = theta_vals, power = delta1) %>%
  ggplot(., aes(x = theta, y = power)) +
  geom_line()+
  xlab(expression(theta)) +
  theme(text = element_text(size = 14))
```

We can use the plot of this power function to talk about the probability of Type I and II error under $\delta_{1}$.

## Comparing power

Consider the second testing procedure $\delta_{2}$: $r(\mathbf{X}) = |\sum_{i=1}^{10} X_{i} - 5| \text{ with rejection region } R = \{r(\mathbf{X}) \geq 4 \}$.

So the power function is $\pi(\theta | \delta_{1}) = P( \sum_{i=1}^{10} X_{i} \in \{0, 1, 9, 10\}|\theta)$.

```{r}
delta2 <- power_fn(theta_vals, n, c(0, 1, 9, 10))
data.frame(theta = theta_vals, delta1, delta2) %>%
  pivot_longer(cols = 2:3, names_to = "procedure", values_to = "power") %>%
  ggplot(.,aes(x = theta, y = power, col = procedure)) +
  geom_line() + 
  scale_color_manual(values = c("black", "purple"))+
  xlab(expression(theta)) +
  theme(text = element_text(size = 14))
  
```

How do the two testing procedures $\delta_{1}$ and $\delta_{2}$ compare with respect to the probabilities of committing a Type I vs. Type II error?
