---
title: "EM algorithm: Extra Credit"
author: "Your name"
date: "2024-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Be sure to set `eval = TRUE` before submitting!

## E-step function

Fill in the remainder of the `e_step` function which takes in the observed phenotype counts `x` (vector of length 3) and allele frequencies `p` (vector of length 3). Using `x` and `p`, the function should return a vector of length six of the expected genotype count for each of the six genotypes.

```{r eval = FALSE}
# x = observed phenotype counts (carbonaria, insularia, typica)
# p = allele probabilities (carbonaria, insularia, typica)

e_step <- function(x, p){
  n_cc <- 
  n_ci <- 
  n_ct <- 
  n_ii <-
  n_it <- 
  n_tt <-
  return()
}
```

## M-step function

Fill in the remainder of the `m_step` function which takes in the observed phenotype counts `x` (vector of length 3) and expected genotype counts `n` (vector of length 6). This function should return a vector of length three of the new estimate of $\boldsymbol{\theta}$.

```{r eval = FALSE}
# x = observed phenotype counts (carbonaria, insularia, typica)
# n = expected genotype counts (CC, CI, CT, II, IT, TT)
m_step <- function(x, n){
  p_c <- 
  p_i <- 
  p_t <- 
  return()
}
```

## Implementation

We have the following observed counts/data of the three phenotypes for the peppered moths: $X = (N_{C} = 85, N_{I} = 196, N_{T} = 341)$.

```{r}
x <- c(85, 196, 341)
```

Now to implement the algorithm, complete the following:

1.  Create a vector `p_old` of length three that initializes the algorithm with the three alleles being equal in frequency.
2.  Create a variable `n_sim` that sets the number of iterations you'd like to run the algorithm for. I recommend you start with something relatively small just to test your code. You can always modify this value later on!
3.  Create a matrix `p_mat` that will hold/store the $\hat{\mathbf{p}}$ estimates from each iteration of the algorithm. Use the `matrix()` function to do this (think about what dimensions you'll need `p_mat` to have). You can fill the matrix with `NA` or `0` values.
4.  Replace the first row of `p_mat` with your initial value/guess for $\mathbf{p}$.
5.  Complete the `for` loop where you iterate between the E and M steps by calling the appropriate functions and passing in the appropriate arguments. You may need to create an additional variable or two.

Then go ahead and run this code chunk (make sure you've run all of the previous chunks first).

```{r eval = FALSE}
# 1. Initialize

# 2. Set number of iterations

# 3. Create p_mat

# 4. Replace first row of p_mat

# 5. for loop to run algorithm

for(i in 2:n_sim){
  # e_step
  
  # m_step
  
  # store estimate in p_mat

}
```

### Assess convergence and state results.

1.  Run your algorithm until convergence. Roughly how many iterations did we need until we hit convergence?
    a.  **Solution:**
2.  What are your maximum likelihood estimates for the frequencies of the three alleles $C$, $I$, and $T$ under the assumption of Hardy-Weinberg?
    a.  **Solution**:
3.  Optional: what are the maximum likelihood estimate for the frequencies of the three alleles $C$, $I$, and $T$ if we don't assume Hardy-Weinberg?
    a.  **Solution**:
