---
title: "MATH 311: Homework 7 (R)"
author: "Your name here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 2

### Part b

Write a function in R which will take a vector $\mathbf{x}$ as input and outputs the value of the skewness $M_{3}(\mathbf{x})$ for this vector.

```{r prob-2b}

```

### Part c

A sample of size 20 was generated from a skewed distribution with unknown CDF $F$. Generate 5000 bootstrap samples from `x`. Obtain approximations of the bias and standard deviation of the sample skewness estimator $M_{3}$ using your bootstrapped statistics.

```{r prob-2c}

```

**Solution:**

## Problem 3

### Part a

Compute the estimate $\hat{e}$ for this sample.

```{r prob-3a}
# you can optionally do this part on paper
```

### Part b

Either answer on paper or type your solution here!

### Part c

Use the `rbinom` function to generate 10000 bootstrap estimates for $(\hat{p}_{c}^{*}, \hat{p}_{v}^{*})$. Then compute bootstrap estimates $\hat{e}^{*}$ based on each of these pairs of bootstrap estimates. 

```{r prob-3c}

```

### Part d

Create a histogram of the bootstrap distribution of $\hat{e}^{*}$ and describe the shape, center, and spread of the distribution.

```{r prob-3d}

```

**Solution:** 

### Part e

Create a 95% bootstrap confidence interval for $e$.

```{r prob-3e}

```

**Solution:** 

## Problem 4

### Part a


```{r prob-4a-fns}

```

Notice the `cache = T` in the chunk headers below. Because this code will take a long time to run, it would be extremely annoying to have to wait a minute for your document to knit every time. We can avoid this by *caching* our results. First we name the chunk (e.g. `prob-4a`). Then when we knit, the executed code from this chunk gets saved in a file. Every time you knit the document again, R simply uses the saved output, if the code in the chunk `prob-4a` is unchanged. Anytime you change the code in the code chunk, upon knitting, R will re-run that code chunk and overwrite the saved output.

```{r prob-4a, cache = T}

```

**Solution:**


### Part b

```{r prob-4b, cache = T}

```

**Solution:** 
