---
title: "R and RStudio review"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
---

\vspace{-2cm}


### Set up

1.  Make a STAT 311 folder (suggested)
2.  Making a new document
    -   File -\> R Markdown... or Quarto Document... (You don't have to name it yet, but do put your name!)
3.  Saving work via save or knitting/rendering
    -   Save to proper location!

    -   HTML or PDF

### Basic coding 

R code in a `.Rmd` or `.qmd` goes in a code chunk. Make comments using the pound symbol

```{r}
# simple arithmetic
1 * 4
```

Ways to run code:

-   Run a specific line of code by clicking anywhere on that line and hidding Cmd + Enter

-   Run all code in a specific code chunk by clicking on green triangle

-   Run all code in previous code chunks (sequentially) but NOT including this code chunk by clicking on gray triangle

-   Render/knit document

```{r}
# create a vector using c() and store as variable 
x <- c(1, 3, 5)

# R is vectorized language
2 * x
```

## Basic plotting

```{r}
xx <- seq(0, 10, 0.5)
plot(xx, 2*xx, 
     type = "l", xlab = "x-axis title", ylab = "y-axis title", main = "Title")
```


### Functions related to named distributions

Suppose $X \sim \text{Binom}(8, 0.75)$.

```{r}
# P(X = 6)
dbinom(6, size = 8, prob = 0.75)

# P(X <= 6)
pbinom(6, size = 8, prob = 0.75)

# What is the median of this distribution? i.e. what is the 50th percentile?
qbinom(0.5, size = 8, prob = 0.75)

# Let's generate 10 realizations from this binomial reproducibly
set.seed(311)
rbinom(10, size = 8, prob = 0.75)

# let's approximate (empirically) the mean of the this binomial using Law of Large Numbers
samps <- rbinom(1000, size = 8, prob = 0.75)
mean(samps)
```

R has these functions for all named distributions (e.g. Poisson `dpois()`, Hypergeometric `dhyper()`, Normal `dnorm()`, etc.) But be careful with parameterizations!

### Optional: ggplot

```{r message = F, fig.height=2.5, fig.width=5}
library(tidyverse) # may need to install
data.frame(x = xx, y = 2*xx) |>
  ggplot(aes(x = x, y = y)) +
  geom_point()
```

